Using best practices for forecasting biodiversity changes in breeding birds

### 2017-07-27 manuscript/draft.Rmd
edits

Developing forecasts for biodiversity is important for conservation, management,
and evaluating how well current models characterize natural systems. While there
are an increasing number of studies forecasting biodiversity, there is little
information available on how well these forecasts work because these studies
fail to follow good practice for making and evaluating forecasts. Most diversity
forecasts do not conduct meaningful evaluations of the forecasts, fail to
account for uncertainty, and do not use time-series data that captures the
actual dynamics being studied. Using hindcasting we made and evaluated XXXX
[number of years * number of models * number of sites] forecasts for breeding
bird species richness at XXX sites throughout North America following best
practices for forecasting. Forecasts were made for up to a decade into the
future. Our results show that while all of the models we evaluated produce
reasonable forecasts, none of them outperforms the simple baseline of long-term
average richness. This means that at present our best forecast for future
richness is simply to assume that it is unchanging. This calls into question the
efficacy of current modeling approaches. Most of the best practices implemented
in this study directly influence the outcome of the forecasts and result in
different conclusions from those made in previous studies. This emphasizes the
importance of following good practice in making ecological forecasts and of
starting to actively make more forecasts in such a way that they can be quickly
evaluated and improved on.

@@

When comparing forecasts for richness across sites all methods performed well
(Figure 2; R^2 XX-XX%). However SDMs (both stacked and joint) and the
macroecological model all failed to successfully forecast the highest-richness
sites resulting in a notable clustering of predicted values near ~60 species and
the poorest performance in the cross-site comparison (R^2 XX-XX%).

While the models generally did well in absolute terms (Figure 2), they tended
not to outperform the "average" baseline from the generalized linear mixed model
(Figure 3A). [[Stats go here]].  The auto-ARIMA was generally the
best-performing non-baseline model, but in many cases (67% of the time), the
best-fitting ARIMA model chosen by the model selection procedure had no
autoregressive, drift, or moving average terms, making it similar to the
"Average" model.

Most models produced confidence intervals that were too narrow, indicating
overconfident predictions (Figure 3B). The random forest-based SDM stack was the
most overconfident model, with only XX% of observations falling inside its 95%
confidence intervals. This stacked SDM's narrow predictive distribution also
caused it to have notably higher deviance (Figure 3C) than the next-worst model,
even though its point estimates were not unusually bad in terms of RMSE (3A). As
discussed elsewhere [@harris2015], this overconfidence is a product of the
assumption in stacked SDMs that errors in the species-level predictions are
independent. Because they avoided this assumption, the GBM-based
"macroecological" model and especially the mistnet JSDM had reasonably
well-calibrated uncertainty estimates (Figure 3B); as a result, their relative
performance was higher in terms of deviance than in terms of RMSE. The "naive"
model was the only model whose confidence intervals were too wide (Figure 3B),
which can be attributed to the rapid rate at which these intervals expand as the
time lag increases (Figure 1).

>>>>

Developing forecasts for biodiversity is important for conservation, management,
and evaluating how well current models characterize natural systems. While there
are an increasing number of studies forecasting biodiversity, there is little
information available on how well these forecasts work because these studies
fail to follow good practice for making and evaluating forecasts. Most diversity
forecasts do not conduct meaningful evaluations of the forecasts, fail to
account for uncertainty, and do not use time-series data that captures the
actual dynamics being studied. Using hindcasting we made and evaluated `r
numbers$n_forecasts` forecasts for breeding bird species richness at `r
numbers$N_predict_sites` sites throughout the continental United States,
following best practices for forecasting. Forecasts were made for up to a decade
after the final training observations. Our results show that while all of the
models we evaluated produce reasonable forecasts, none of them consistently
outperforms the simple baseline of long-term average richness. This means that
at present, our best forecast for future richness is simply to assume that it is
unchanging. This calls into question the efficacy of current modeling
approaches. Most of the best practices implemented in this study directly
influence the outcome of the forecasts and result in different conclusions from
those made in previous studies. This emphasizes the importance of following good
practice in making ecological forecasts and of starting to actively make more
forecasts in such a way that they can be quickly evaluated and improved on.

@@

When comparing forecasts for richness across sites all methods performed well
(Figure 2; all $R^2 > 0.5$). However SDMs (both stacked and joint) and the
macroecological model all failed to successfully forecast the highest-richness
sites resulting in a notable clustering of predicted values near ~60 species and
the poorest performance ($R^2$ `r numbers$env_R2s`, versus XX-XX for the
within-site methods).

While the models generally did well in absolute terms (Figure 2), none tended to
outperform the "average" baseline from the generalized linear mixed model
(Figure 3A). [[Stats go here]].  The auto-ARIMA was generally the
best-performing non-baseline model, but in many cases (67% of the time), the
best-fitting ARIMA model chosen by the model selection procedure had no
autoregressive, drift, or moving average terms, making it similar to the
"Average" model.

Most models produced confidence intervals that were too narrow, indicating
overconfident predictions (Figure 3B). The random forest-based SDM stack was the
most overconfident model, with only `r numbers$rf_coverage_pct`% of observations
falling inside its 95% confidence intervals. This stacked SDM's narrow
predictive distribution also caused it to have notably higher deviance (Figure
3C) than the next-worst model, even though its point estimates were not
unusually bad in terms of RMSE (3A). As discussed elsewhere [@harris2015], this
overconfidence is a product of the assumption in stacked SDMs that errors in the
species-level predictions are independent. Because they avoided this assumption,
the GBM-based "macroecological" model and especially the mistnet JSDM had
reasonably well-calibrated uncertainty estimates (Figure 3B); as a result, their
relative performance was higher in terms of deviance than in terms of RMSE. The
"naive" model was the only model whose confidence intervals were too wide
(Figure 3B), which can be attributed to the rapid rate at which these intervals
expand as the time lag increases (Figure 1).
